
# README -----------------------------------------------------------------------

Workflow for processing PRS probe data at Desert Fertilization sampling
locations. The format of data provided from WesternAg can vary just enough each
iteration that functionalizing this workflow has not worked well, and the
sampling is infrequent enough that it is easier to simply template the workflow
of a previous run for new runs.


# config -----------------------------------------------------------------------

```{r config, eval=FALSE}

source("config.R")

```

# detection limits -------------------------------------------------------------

Unique analyte names when imported with read_excel() and after renaming `Total
N` to `Total-N`:

```{r analytes, eval=FALSE, echo=TRUE, message=TRUE}

analytes <- c(
  "Total-N",
  "NO3-N",
  "NH4-N",
  "Ca",
  "Mg",
  "K",
  "P",
  "Fe",
  "Mn",
  "Cu",
  "Zn",
  "B",
  "S",
  "Pb",
  "Al",
  "Cd"
)

```

```{r detection-limits, eval=TRUE}

two <- c(
  "Total-N",
  "NO3-N",
  "NH4-N",
  "Ca",
  "S"
)

four <- c(
  "Mg",
  "K"
)

two_tenths <- c(
  "P",
  "Mn",
  "Cu",
  "Zn",
  "B",
  "Pb",
  "Cd"
)

four_tenths <- c(
  "Fe",
  "Al"
)

```

# import

```{r import, eval=TRUE}

prs_upload <- readxl::read_excel(
  path = "~/Dropbox/development/PRS/2021/Nutrient Supply Rate Data_ Project 2198_ Quincy Stewart.xlsx",
  skip = 5
)

# check for reasonable (< 180 d) deployment period
prs_upload |>
  pointblank::col_vals_lt(
    columns       = vars(date_range),
    value         = 180,
    na_pass       = TRUE,
    preconditions = function(x) { dplyr::mutate(x, date_range = lubridate::ymd(x$`Retrieval Date`) - lubridate::ymd(x$`Burial Date`))},
    actions       = pointblank::stop_on_fail()
  )

```

# processing

```{r processing, eval=TRUE}

new_prs <- prs_upload |>
  dplyr::filter(!is.na(`Sample ID`)) |>
  dplyr::rename(`Total-N` = `Total N`) |>
  tidyr::pivot_longer(
    cols      = `Total-N`:last_col(),
    names_to  = "id",
    values_to = "result"
    ) |>
  # gather(id, result, `Total-N`:last_col()) |> # stack
  dplyr::mutate(
    plotid = as.numeric(gsub("[[:alpha:]]", "", `Sample ID`)),
    location = dplyr::case_when(
      grepl("a", `Sample ID`, ignore.case = TRUE) ~ "under plant",
      grepl("b", `Sample ID`, ignore.case = TRUE) ~ "between plant",
      plotid > 75 ~ "BLANK",
      TRUE ~ NA_character_
      ),
    flag = dplyr::case_when(
      result <= 2.0 & id %in% two ~ "below detection limit",
      result <= 4.0 & id %in% four ~ "below detection limit",
      result <= 0.2 & id %in% two_tenths ~ "below detection limit",
      result <= 0.4 & id %in% four_tenths ~ "below detection limit",
      TRUE ~ NA_character_
    )
  )

```

# add data to the database -----------------------------------------------------

## temporary table to postgres

```{r temporary-table-to-pg, eval=TRUE}

# write new data data to a temporary table

if (dbExistsTable(pg, c("urbancndep", "new_prs"))) {

  dbRemoveTable(pg, c("urbancndep", "new_prs"))

}

dbWriteTable(pg, c("urbancndep", "new_prs"), value = new_prs, row.names = F)

```

## format temporary table

Need to alter database fields namely because dbWriteTable creates a date field
with time zones. Changing the Wal ID to integer may or may not be required
depending on how WesternAg provides the data.

Note that WesternAg often uses slightly different column names so these may
need to be tweaked for each run.

```{r format-table, eval=TRUE}

DBI::dbWithTransaction(pg,
  DBI::dbExecute(pg,'
    ALTER TABLE urbancndep.new_prs
    ALTER COLUMN "Burial Date" TYPE date USING ("Burial Date"::date),
    ALTER COLUMN "Retrieval Date" TYPE date USING ("Retrieval Date"::date),
    ALTER COLUMN "WAL #" TYPE integer USING ("WAL #"::integer);'
  )
)

```

## insert data from temporary table to prs_analysis

As with formatting (above), WesternAg often uses slightly different column
names so the insert statement may need to be tweaked for each run.

```{r write-to-disk, eval=TRUE}

DBI::dbWithTransaction(pg,
  DBI::dbExecute(pg,'
    INSERT INTO urbancndep.prs_analysis
    (
      wal_id,
      plot_id,
      start_date,
      end_date,
      analyte,
      final_value,
      flag,
      location_within_plot,
      num_cation_probes,
      num_anion_probes,
      notes
    )
    (
      SELECT
      "WAL #",
      plotid,
      "Burial Date",
      "Retrieval Date",
      id,
      result,
      flag,
      location,
      "# Cation",
      "# Anion",
      "Notes"
      FROM
      urbancndep.new_prs
      );')
  )

```

## house keeping

```{r housekeeping, eval=TRUE}

dbRemoveTable(pg, c("urbancndep", "new_prs"))

```

